---
title: "NYPD assignment"
output: pdf_document
date: "2023-03-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```
```{r  echo = FALSE , message = FALSE}
library(tidyverse)
library(lubridate)

```

## NYPD Shooting analysis

This document uses the dataset NYPD Shooting Incident Data from Data.gov. It contains a list of every shooting incident that occurred in NYC from 2006 through the end of 2022.

 <https://catalog.data.gov/dataset/nypd-shooting-incident-data-historic>.


```{r import}
# read the csv
nypd <- read_csv("https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD")
```
\newpage 
# Creating datasets for analysing the data
```{r analysis}
#dataset by BORO
nypd_boro <-nypd %>% group_by(BORO, LOCATION_DESC) %>%   
  summarize(cases = n()) %>% 
  select(BORO, LOCATION_DESC, cases) %>% 
  ungroup()

#dataset by location
nypd_loc <-nypd %>%filter(!is.na(LOCATION_DESC))

#renaming some of the values 
nypd_loc$LOCATION_DESC <- recode(nypd_loc$LOCATION_DESC,
"MULTI DWELL - APT BUILD" = "MULTI DWELL",
"MULTI DWELL - PUBLIC HOUS" = "MULTI DWELL",
"BAR/NIGHT CLUB" = "CLUB",
"BEAUTY/NAIL SALON" = "SALON",
"COMMERCIAL BLDG" = "COMM BLDG",
"GROCERY/BODEGA" = "GROCERY",
"PVT HOUSE" = "HOUSE",
"RESTAURANT/DINER" = "REST/DINER")

# Case count by location 
nypd_loc <- nypd_loc  %>%filter(!is.na(LOCATION_DESC))%>%
group_by(VIC_SEX, LOCATION_DESC) %>%
summarize(cases = n()) %>%
select( VIC_SEX, LOCATION_DESC, cases) %>%
ungroup()


# Dataset by victim's sex in each Boro
nypd_vic_sex <-nypd %>% group_by(VIC_SEX, BORO) %>% 
  summarize(cases = n()) %>% 
  select( VIC_SEX, BORO, cases) %>% 
  ungroup()


#dataset by Year
nypd_date <-nypd %>% 
  mutate(OCCUR_YEAR = year(as.Date(OCCUR_DATE, "%m/%d/%Y"))) %>% 
  group_by(VIC_SEX, OCCUR_YEAR) %>% 
  summarize(cases = n()) %>% 
  select( VIC_SEX, OCCUR_YEAR, cases) %>% 
  ungroup()
```
\newpage

# Location with more than 100 cases 
```{r viewbyloc}
#selecting the location with more than 100 cases

nypd_loc %>% filter(cases > 100) %>% 
  ggplot(aes(x= LOCATION_DESC, y = cases)) +   
  geom_point(aes(color = VIC_SEX)) +
  labs(title = "Location with more than 100 cases")
```
\newpage

# View the number of cases in each year starting from 2006 until 2022
```{r viewbyyear}

nypd_date %>% ggplot(aes(x= OCCUR_YEAR, y = cases,color = VIC_SEX)) +    
  geom_point() +
  labs(title = "Cases by Year")
``` 

\newpage

# Data Model
Generating a model to determine the percentage of cases for each boro against the total number of cases 

```{r model}
# using the vic_sex dataset to create a model 
nypd_vic_sex <- nypd_vic_sex %>% 
  mutate(total_count = sum(cases), 
         case_pct = cases/total_count * 100)

# Determine the total number of cases for each BORO
nypd_vic_sex <- nypd_vic_sex  %>% group_by(BORO) %>% 
  mutate(boro_count = sum(cases), boro_pct = cases/boro_count * 100) %>% 
  ungroup()

#creating a model 
mod <- lm(boro_pct ~ case_pct , data = nypd_vic_sex)

#Summary of the model 
summary(mod)

#creating the prediction for the model 
nypd_mod_totals <- nypd_vic_sex %>%  mutate(pred = predict(mod))

# Plotting the model to visualize the data 
nypd_mod_totals %>% ggplot() +
  geom_point(aes(x=case_pct, y = boro_pct), color = "green" ) + 
  geom_point(aes(x=case_pct, y=pred), color = "brown") +
  labs(title = "Data Model")
```
\newpage

# Bias

For me the whole of NYC looks unsafe. I'm scared to even visit the city. NYC schools or 
universities have been safer unlike the other places in the US. 
USA needs to control gun violence all over the country!!

While working on this data it shows that crimes committed on males is more when 
compared to females.
Starting from 2015 the crime on men was declining and increased drastically again in 2020.
If the state department was implementing certain measures between 2015 - 2020 that 
reduced the crimes during that period, it should be re-enforced after the pandemic.